---
title: "Penguin Flipper Lengths ANOVA"
author: "Tess Hooper"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Load packages and read in the penguins.csv data

Packages: tidyverse, car, onewaytests

```{r}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")
```

Our question: Is there a significant differnce in mean flipper lengths for female Chinstrap, Gentoo, and Adelie Penguins near Palmer Archipelago #single factor=species, 3 levels=type of species, random variable= flipper length

Maybe I can use a one-way ANOVA...
- Independant observations
- Equal variance (Leven's test)
- Normality (histograms, qqplots)


```{r}

counts <- with(penguins, table(species, sex)) #with the dataset penguins, make a data table that counts species and sex

penguins2 <- penguins %>% 
  mutate( #create a new column based on following paramters - column called common_name, with values where species with name of "Chinstrap penguin", call it "chinstrap"
    common_name = 
      case_when(
        species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
        species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
        species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
  ) %>% 
  
  select (-species) %>% 
  filter(sex== "FEMALE")
```



