---
title: "Penguin Flipper Lengths ANOVA"
author: "Tess Hooper"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Load packages and read in the penguins.csv data

Packages: tidyverse, car, onewaytests

```{r}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")
```

Our question: Is there a significant differnce in mean flipper lengths for female Chinstrap, Gentoo, and Adelie Penguins near Palmer Archipelago #single factor=species, 3 levels=type of species, random variable= flipper length

Maybe I can use a one-way ANOVA...
- Independant observations
- Equal variance (Leven's test)
- Normality (histograms, qqplots)


```{r}

counts <- with(penguins, table(species, sex)) #with the dataset penguins, make a data table that counts species and sex

penguins2 <- penguins %>% 
  mutate( #create a new column based on following paramters - column called common_name, with values where species with name of "Chinstrap penguin", call it "chinstrap"
    common_name = 
      case_when(
        species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
        species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
        species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
  ) %>% 
  
  select (-species) %>% 
  filter(sex== "FEMALE")
```


Exploratory graphs and Levene's test for equal variances

```{r}

hists <- ggplot(penguins2, aes(x=flipper_length))+
  geom_histogram(aes(fill = common_name))+
  facet_wrap(~common_name, scale="free")

hists


qqplot <-ggplot(penguins2, aes(sample=flipper_length))+
  geom_qq(aes(color = common_name))+
   facet_wrap(~common_name, scale="free")

qqplot

# Based on qq-plots and histograms, observations for each group appear normally distributed. Also, n > 30 for all groups, so means will be normally distributed by CLT.
```


Evaluate variances and perfrom a Levene's Test for equal variances

```{r}

# If the largest sample variances is < 4x greater than the smallest sample variance, then these are "close enough" to consider variances equal

variances <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = var(flipper_length)
  )

variances #variance units will be in mm^2 ... The largest variance (33.1) is definitely not more than 4x greater than the smallest variance (15.2)

# Levene's Test
# leveneTest function in the 'car' package


levene_penguin <- leveneTest(flipper_length ~ common_name, data= penguins2)

levene_penguin

# H0: No difference in variances (variances are equal)
# HA: Variances are NOT equal

# There is no significant difference in variances across species, 
```



One-way ANOVA for penguin flipper lengths by species (common_name)

```{r}

# H0: THe means across all groups are equal
# HA: At least two group means differ significantly 

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)
summary(penguin_aov)


# Post-hoc testing by Tukey's HSD


#H 0: No significant difference in pairwise means
# HA: There IS a significant difference in pairwise means

penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph

# All p-values for each species is very low, meaning there are differences among all of them
```

Mean flipper lengths of female penguins differed significantly for all species studied (one-way anova, F(df) = statistic, p < 0.001, $\alpha$ = 0.05 with post-hoc Tukey's HSD, $\alpha$ = 0.05).



